---
title: "Data_visualization"
author: "Ian Greaves"
date: "8 August 2018"
output: html_document
---

#Data:
For this task I've decided to use the legumes data set.  The data is as follows

'Rhizobes', i.e. bacteria living on plant roots, can increase plant growth. This dataset investigates the effect of different rhizobial strains on legume growth. Variable are as follows:

**Block:** independent experimental 'block'

**Plant:** plant species (Ad: Acacia dealbata; Am: Acacia mearnsii; Dm: Daviesia mimosoides; Gc: Glycine clandestina; Hv: Hardenbergia violacea; Oe: Oxylobium ellipticum)

**Strain:** rhizobial strain (note: values are descriptive strain identifiers, not quantitative measures) 20 strains in total

**Weight:** plant weight at conclusion of experiment.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

## Setting up environment
Loading library dplyr

```{r}
library(tidyverse)
```

##Adding in data
```{r}
Legumes <- read_csv(file = "data/legume_weights_rhizobia_20str.csv")
Legumes
```

##Make Data tidy
Will make the data tidy by combining the first 3 lists to make a unique identifier.  Will retain all the lists for visualization
```{r}
Legumes1 <- Legumes %>%
  unite(Identifier,Block, Plant, Strain, sep = "_")
Legumes1
```
That didn't work I will use mutate
```{r}
Legumes = mutate(Legumes,
   Identifier = paste(Block, Plant, Strain, sep = '_'))
Legumes <- Legumes[,c(5,1:4)]
Legumes
```


## Initial Analysis
Look at the overall data.
```{r}
ggplot(data = Legumes,
       aes(x = as.factor(Strain), y = weight, colour = Plant)) +
  geom_point() +
  labs(
    x = "Strain",
    y = "Weight",
    title = "Figure 1 - All data looking at strains",
    colour = "Plant"
  ) +
  theme(
    panel.grid=element_blank(),
    axis.text.x = element_text(angle=90),
    panel.background = element_rect(fill = "white",
                                    colour = "black")
    )
```  
Change the grouping to PLant and see if the scatter plot can quickly show which strain was best for each plant
```{r}
ggplot(data = Legumes,
       aes(x = Plant, y = weight, colour = as.factor(Strain))) +
  geom_point() +
  labs(
    x = "Plant",
    y = "Weight",
    title = "Figure 1 - All data looking at plants",
    colour = "Strain"
  ) +
  theme(
    panel.grid=element_blank(),
    axis.text.x = element_text(angle=90),
    panel.background = element_rect(fill = "white",
                                    colour = "black")
    )
```


Will go back to previous graph and facet by Plant
```{r}
ggplot(data = Legumes,
       aes(x = as.factor(Strain), y = weight)) +
  geom_point() +
  facet_wrap(~ Plant) +
  labs(
    x = "Strain",
    y = "Weight",
    title = "Figure 1 - Separating plants",
    colour = "Plant"
  ) +
  theme(
    panel.grid=element_blank(),
    axis.text.x = element_text(angle=90),
    panel.background = element_rect(fill = "white",
                                    colour = "black")
    )
```
###Conclusion
What I know

What I dont know

Check to see if jitter helps the data
```{r}
ggplot(data = Legumes,
       aes(x = as.factor(Strain), y = weight)) +
  geom_jitter() +
  facet_wrap(~ Plant) +
  labs(
    x = "Strain",
    y = "Weight",
    title = "Figure 1 - Jitter",
    colour = "Plant"
  ) +
  theme(
    panel.grid=element_blank(),
    axis.text.x = element_text(angle=90),
    panel.background = element_rect(fill = "white",
                                    colour = "black")
    )
```

Would like to look and see if there is any variation between blocks
```{r}
ggplot(data = Legumes,
       aes(x = Plant, y = weight, colour = Block)) +
  geom_point(size =1, alpha = 0.5) +
  facet_wrap(~ Strain) +
  labs(
    x = "Plant",
    y = "Weight",
    title = "Figure 1 - Looking at Block variation",
    colour = "Block"
  ) +
  theme(
    panel.grid=element_blank(),
    axis.text.x = element_text(angle=90),
    panel.background = element_rect(fill = "white",
                                    colour = "black")
    )
```
Can't see any obvious differences between blocks

##Stuff NOT to do

To get a feel how each plant has done with each strain I/m giong to use dplyr group by to group by plant and strain and use mean to xummarize the weights.  Data contains NA's so will remove them.

```{r}
Legumes_Av <-Legumes %>% 
  drop_na() %>% 
    group_by(Plant, Strain) %>% 
    summarise(mean_weight=mean(weight), n=n(), sd = sd(weight),se=sd/sqrt(n))

Legumes_Av
```



```{r}
ggplot(data = Legumes,
       aes(x = as.factor(Strain), y = weight)) +
  geom_boxplot() +
  facet_wrap(~ Plant) +
  labs(
    x = "Strain",
    y = "Weight",
    title = "Figure 1 - All data",
    colour = "Plant"
  ) +
  theme(
    panel.grid=element_blank(),
    axis.text.x = element_text(angle=90),
    panel.background = element_rect(fill = "white",
                                    colour = "black")
    )
```




























